---
title: "adeshpa1_11"
author: "Amruta Deshpande"
date: "11/22/2020"
output: html_document
---
# Question:

AP is a shipping service that guarantees overnight delivery of packages in the continental US.
The company has various hubs at major cities and airports across the country. Packages are
received at hubs, and then shipped to intermediate hubs or to their final destination.
The manager of the AP hub in Cleveland is concerned about labor costs, and is interested in
determining the most effective way to schedule workers. The hub operates seven days a week,
and the number of packages it handles varies from one day to another. The table below provides
an estimate of the number of workers needed each day of the week.

Day of the Week Workers Required

Sunday 18
Monday 27
Tuesday 22
Wednesday 26
Thursday 25
Friday 21
Saturday 19

Package handlers at AP are guaranteed a five-day work week with two consecutive days off. The
base wage for the handlers is $750 per week. Workers working on Saturday or Sunday receive an
additional $25 per day. The possible shifts and salaries for package handlers are:

Shift Days Off Wage

1 Sunday and Monday $775
2 Monday and Tuesday $800
3 Tuesday and Wednesday $800
4 Wednesday and Thursday $800
5 Thursday and Friday $800
6 Friday and Saturday $775
7 Saturday and Sunday $750


The manager wants to keep the total wage expenses as low as possible while ensuring that there
are sufficient number of workers available each day. Formulate and solve the problem. What was
the total cost? How many workers are available each day?

# Solution:

Let us consider the decision variables as following:   
X1= the number of workers assigned to shift 1  
X2= the number of workers assigned to shift 2  
X3= the number of workers assigned to shift 3  
X4= the number of workers assigned to shift 4  
X5= the number of workers assigned to shift 5  
X6= the number of workers assigned to shift 6  
X7= the number of workers assigned to shift 7  


Loading library "lpSolveAPI" and we have 7 decision variables and 7 constraints

```{r}

library(lpSolveAPI) 
 
AP_lp <- make.lp(7, 7)

```

Setting the objective function 

```{r}

set.objfn(AP_lp, c(775, 800, 800, 800, 800, 775, 750))

lp.control(AP_lp,sense='min')

```
Setting LHS and RHS

```{r}

set.row(AP_lp, 1, c(0, 1, 1, 1, 1, 1, 0), indices = c(1, 2, 3, 4, 5, 6, 7))
set.row(AP_lp, 2, c(0, 0, 1, 1, 1, 1, 1), indices = c(1, 2, 3, 4, 5, 6, 7))
set.row(AP_lp, 3, c(1, 0, 0, 1, 1, 1, 1), indices = c(1, 2, 3, 4, 5, 6, 7))
set.row(AP_lp, 4, c(1, 1, 0, 0, 1, 1, 1), indices = c(1, 2, 3, 4, 5, 6, 7))
set.row(AP_lp, 5, c(1, 1, 1, 0, 0, 1, 1), indices = c(1, 2, 3, 4, 5, 6, 7))
set.row(AP_lp, 6, c(1, 1, 1, 1, 0, 0, 1), indices = c(1, 2, 3, 4, 5, 6, 7))
set.row(AP_lp, 7, c(1, 1, 1, 1, 1, 0, 0), indices = c(1, 2, 3, 4, 5, 6, 7))


rhs <- c(18, 27, 22, 26, 25, 21, 19)

set.rhs(AP_lp, rhs)

```

Setting constraint and variables

```{r}

set.constr.type(AP_lp, c(">=", ">=", ">=", ">=", ">=", ">=", ">="))

set.bounds(AP_lp, lower = rep(0, 7))

set.type(AP_lp, 1:7,"integer")

```


```{r}

lp.rownames <- c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")

lp.colnames <- c("Shift1", "Shift2", "Shift3","Shift4", "Shift5", "Shift6","Shift7")

dimnames(AP_lp) <- list(lp.rownames, lp.colnames)

AP_lp

```

Solving and saving

```{r}

write.lp(AP_lp, filename = "AP.lp", type = "lp")

solve(AP_lp)

```

Getting objective, variables and constraints

```{r}

get.objective(AP_lp)

get.variables(AP_lp)

get.constraints(AP_lp)

```

LP formulation 

```{r}

A <- read.lp("AP.lp")

solve(A)

A

```

Getting objective, variables and constraints

```{r}

get.objective(A)

get.variables(A)

B <- data.table::data.table(Decision_Variables=lp.rownames<-c("x1","x2","x3","x4","x5","x6","x7"),values=get.variables(A))

B

get.constraints(A)

C <- data.table::data.table(Days_of_week=lp.rownames<-c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday"),Workers_available=get.constraints(A))

C

```
The number of workers available are in the table while total cost is 25675.