---
title: "Assignment_4"
author: "Amruta Deshpande"
date: "10/1/2020"
output:
  word_document: default
  pdf_document: default
  html_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# * Question:

#The Weigelt Corporation has three branch plants with excess production capacity. Fortunately,
#the corporation has a new product ready to begin production, and all three plants have this
#capability, so some of the excess capacity can be used in this way. This product can be made in
#three sizes--large, medium, and small--that yield a net unit profit of $420, $360, and $300,
#respectively. Plants 1, 2, and 3 have the excess capacity to produce 750, 900, and 450 units per
#day of this product, respectively, regardless of the size or combination of sizes involved.
#The amount of available in-process storage space also imposes a limitation on the production
#rates of the new product. Plants 1, 2, and 3 have 13,000, 12,000, and 5,000 square feet,
#respectively, of in-process storage space available for a day's production of this product. Each
#unit of the large, medium, and small sizes produced per day requires 20, 15, and 12 square feet,
#respectively.
#Sales forecasts indicate that if available, 900, 1,200, and 750 units of the large, medium, and
#small sizes, respectively, would be sold per day.
#At each plant, some employees will need to be laid off unless most of the plant’s excess
#production capacity can be used to produce the new product. To avoid layoffs if possible,
#management has decided that the plants should use the same percentage of their excess capacity
#to produce the new product.
#Management wishes to know how much of each of the sizes should be produced by each of the
#plants to maximize profit.
#Solve the problem using lpsolve, or any other equivalent library in R.

# * Solution:

#Loading library "lpSolveAPI"

```{r}

library(lpSolveAPI)

```

#Making a lp object with 9 constraints and decision variables

```{r}

lprec <- make.lp(9, 9)

```

#Creating Objective function

```{r}

set.objfn(lprec,c(420, 360, 300, 420, 360, 300, 420, 360, 300))

```

#As default is minimization problem, changing direction to maximization

```{r}

lp.control(lprec, sense='max')

```

#Setting a column and specifying the numeric vector and indices

```{r}

set.row(lprec, 1, c(1, 1, 1), indices = c(1, 2, 3))
set.row(lprec, 2, c(1, 1, 1), indices = c(4, 5, 6))
set.row(lprec, 3, c(1, 1, 1), indices = c(7, 8, 9))
set.row(lprec, 4, c(20, 15, 12), indices = c(1, 2, 3))
set.row(lprec, 5, c(20, 15, 12), indices = c(4, 5, 6))
set.row(lprec, 6, c(20, 15, 12), indices = c(7, 8, 9))
set.row(lprec, 7, c(1, 1, 1), indices = c(1, 2, 3))
set.row(lprec, 8, c(1, 1, 1), indices = c(4, 5, 6))
set.row(lprec, 9, c(1, 1, 1), indices = c(7, 8, 9))

```

#Listing and setting all the right hand sides

```{r}

rhs <- c(750, 900, 450, 13000, 12000, 5000, 900, 1200, 750)

set.rhs(lprec, rhs)

```

#Setting constraints types 

```{r}

set.constr.type(lprec, rep("<=", 9))

```


#Printing the LP model object "lprec" and saving the model to a file. 

```{r}

lprec

write.lp(lprec, filename="WC.lp", type="lp")

```

#Solving the LP problem 

```{r}

solve(lprec)

```

#The output above indicates that it was a successful solution

#Now getting the values of objective function, variables and constraints

```{r}

get.objective(lprec)

get.variables(lprec)

get.constraints(lprec)

```

#The solution shows that the profit is 710000, with the variables XL1 as 350, XM1 as 400, and so on for all the sizes and plants. 


# Considering another method, here we will use the "lp" format. In this method the problem formulation is done in a text file and then input to R. 

#Lets create a text file that contains the problem formulation. Here the text file is save as "Weigelt.lp". 

#Reading the LP model formulation from Weigelt text file. 

```{r}

lprec <- read.lp("WC.lp")

```

#Printing the lprec

```{r}

lprec

```

#Solving the LP problem 

```{r}

solve(lprec)

```


#Getting the values of objective function, variables and constraints

```{r}

get.objective(lprec)

get.variables(lprec)

get.constraints(lprec)

```

#Hence, by solving with both the methods; the profit is 710000 with the Plant1 producing 350 Large size product. While Plant2 producing 500 Medium products and so on.